---
title: "GEOG891_Lab1_Yoo"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message = F)
```

## Practice

## Load Packages

```{r}
library(tidyverse)
library(sp)
library(ggplot2)
library(hrbrthemes)
```

## Load Data

```{r}
p.counties <- "../data/CBW/County_Boundaries.shp"
p.stations <- "../data/CBW/Non-Tidal_Water_Quality_Monitoring_Stations_in_the_Chesapeake_Bay.shp"
d.counties <- sf::read_sf(p.counties)
d.stations <- sf::read_sf(p.stations)
glimpse(d.counties)
glimpse(d.stations)
```

## Select

```{r}
d.counties %>% dplyr::select(GEOID10, ALAND10) %>% head()
d.counties %>% dplyr::select(GEOID10:CLASSFP10) %>% head()
d.counties %>% dplyr::select(-(GEOID10:CLASSFP10)) %>% head()
d.counties %>% dplyr::select(starts_with("C"))
```

## Group

```{r}
d.counties %>% group_by(STATEFP10) %>% mutate(stateLandArea = sum(ALAND10))
d.counties %>% as_tibble() %>% dplyr::select(-geometry) %>% group_by(STATEFP10) %>% summarise(stateLandArea = sum(ALAND10))
```

## Plots

```{r}
d.counties %>% ggplot(., aes(x = as.factor(STATEFP10), y = ALAND10)) + geom_boxplot(aes(fill = STATEFP10))
d.counties %>% ggplot(., aes(x = ALAND10)) + geom_histogram(aes(fill = STATEFP10)) + labs(title = "not the most useful plot, but you get the idea")
```

## CRS

```{r}
d.counties %>% sf::st_crs()
d.stations %>% sf::st_crs()
d.counties %>% sf::st_crs() == d.stations %>% sf::st_crs()
```

```{r}
del.counties <- d.counties %>% dplyr::filter(STATEFP10 == 10)
del.stations <- sf::st_intersection(d.stations, del.counties)
glimpse(del.stations)
plot(del.stations)
del.counties %>% sf::st_area()
```

## Tasks
### Task 1: Bacis Data Manipulation
#### 1.1 For each county, calculate its land area as percentage of the total area (land + water) for that state.

```{r}
perland <- d.counties %>% as_tibble() %>% dplyr::select(-geometry) %>% mutate(Landperc = ALAND10/(ALAND10+AWATER10)*100)
perland[, c('NAME10', 'Landperc')]
```

#### 1.2 For each state, find the county that has the largest proportion of its land as water (water area / total area)

```{r}
landaswater <- d.counties %>% as_tibble() %>% dplyr::select(-geometry) %>% group_by(STATEFP10) %>% summarise(county = NAME10, LargestWaterPerc1 = max(ALAND10/(ALAND10+AWATER10)*100))
landaswater
```

#### 1.3 Count the number of counties in each state

```{r}
d.counties %>% group_by(STATEFP10) %>% mutate(count = n())
d.counties %>% as_tibble() %>% dplyr::select(-geometry) %>% group_by(STATEFP10) %>% summarise(count = n())
```

#### 1.4 Which station has the shortest name (STATION_NA) in the study area?

```{r}
shortChar <- aggregate(d.stations$STATION_NA, list(d.stations$STATION_NA),
    function(s) min(nchar(as.character(s))))
head(shortChar, 1)
```

### Task 2: Plotting attribute data
#### 2.1 Make a scatterplot showing the relationship between land area and water area for each county. Color each point using the state variable

```{r}
require(scales)
p1 <- ggplot(d.counties, aes(x=ALAND10, y=AWATER10, color=STATEFP10)) + 
    geom_point(size=5) + labs(x= "Land Area",
                              y="Water Area",
                              title = "Relationship Between Land Area & Water Area for Each County",
                              colour="State FIPS 10")
    p1 + scale_x_continuous(labels = comma) + scale_y_continuous(labels = comma)
```

#### 2.2 Make a histogram of drainage area (Drainage_A) for all monitoring stations

```{r}
p2 <- ggplot(d.stations, aes(x=Drainage_A)) + 
    geom_histogram() + labs(x= "Drainage Area", title = "Histogram of Drainage Area")
p2
```

#### 2.3 Make a similar histogram, this time of drainage area (Drainage_A) for all monitoring stations. Color each point using the state variable

```{r}
x <- d.stations$STATION_NA
substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}
x1 <- d.stations %>% mutate(station_State = substrRight(x, 2))
x1 %>% as_tibble() %>% dplyr::select(-geometry) %>% group_by(station_State)
p3 <- ggplot(x1, aes(x=Drainage_A, fill=station_State)) + 
    geom_histogram() + labs(x= "Drainage Area", title = "Histogram of Drainage Area")
p3
```

### Task 3: Write a function
#### 3.1 Write a function that does the following:
##### A. accepts a vector of arbitrary numbers, calculates the mean, median, maximum, and minimum of the vector
##### B. Sorts the vector
##### C. returns a list of those values from A and the sorted vector from B
##### D. the function should only work with numeric values and print an error message if any other data type are found
##### Test it with the following vectors
##### c(1, 0, -1), c(10, 100, 1000), c(.1, .001, 1e8), c("a", "b", "c")

```{r}
task3 <- function(...){
  if (is.numeric(...)){
  list <- c(...)
  mean <- mean(list)
  med <- median(list)
  max <- max(list)
  min <- min(list)
  results <- c(mean, med, max, min)
  sort_list <- sort(list)
  cat("List:", sort_list, '\n')
  return(cat('Mean / Med / Max / Min:',results, '\n'))
  } else {
    print("Error: Please insert numeric values")
  }
}

task3(c(1, 0, -1))
task3(c(10, 100, 1000))
task3(c(.1, .001, 1e8))
task3(c("a", "b", "c"))
```

### Task 4: (slightly) more complex spatial analysis.
#### 4.1 Calculate the number of monitoring stations in each state

```{r}
sf::sf_use_s2(F)
all.stations <- sf::st_intersection(d.stations, d.counties)
all.stations %>% group_by(STATEFP10) %>% mutate(count = n())
all.stations %>% as_tibble() %>% dplyr::select(-geometry) %>% group_by(STATEFP10) %>% summarise(count = n())
```

#### 4.2 Calculate the average size of counties in New York (that are also in this study area)

```{r}
x <- all.stations$STATION_NA
substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}
all.stations <- all.stations %>% mutate(station_State = substrRight(x, 2))
ny.counties <- all.stations %>% dplyr::filter(station_State == 'NY')
ny.counties %>% as_tibble() %>% dplyr::select(-geometry) %>% group_by(station_State) %>% summarise(avg_size_NY = mean(ALAND10))
```

#### 4.3 Calculate which state has monitoring stations with the greatest average drainage area (Drainage_A)

```{r}
highdrainage <- all.stations %>% as_tibble() %>% dplyr::select(-geometry) %>% group_by(station_State) %>% summarise(great_drainage = max(Drainage_A))
highdrainage
s.highdrainage <- arrange(highdrainage, desc(great_drainage))
head(s.highdrainage, 1)
```

## Questions
### 1. In using the intersection functions, are the following two statements equivalent? If not, explain how.
#### Be sure to think about BOTH the spatial data structures AND the attribute data. Would your answer be different if we were using different types of data?
##### sf::st_intersection(d.stations, del.counties)
##### sf::st_intersection(del.counties, d.stations)
<br>
I think the statements are equivalent. When I checked attribute data for both orders, I found that objects and variables in the attribute data were the same. The only difference was that the attribute data showed the variables in the first input (x) first and then the second input (y).
<br>
In addition, I found this information on the web.
<br>
Introduction to Spatial Data Programming with R - Michael Dorman
<br>
(https://geobgu.xyz/r/geometric-operations-with-vector-layers.html#introduction-5)
<br>
-	The order of two inputs (d.stations, del.counties) does not matter because st_intersection returns all intersections of the geometries from x,y inputs.
-	There are some other functions that order does not matter (e.x., st_sym_difference and st_union)
-	Order does not matter because the three functions above are symmetrical
-	Order matters when implementing nonsymmetrical functions such as st_difference. This function is not symmetrical because it returns the area of x, minus the intersection of x and y
<br>
Therefore, the order matters depend on the function, not depends on the types of data. When implementing st_intersenction, the result will be the same regardless of order.

```{r}
test1 <- sf::st_intersection(d.stations, del.counties)
test2 <- sf::st_intersection(del.counties, d.stations)
test1
test2
plot(test1)
plot(test2)
```

### 2. What did you find challenging in this lab? What was new?
<br>
Question 1.4 - At first, I was trying this code "d.stations %>% mutate(length = min(nchar(STATION_NA)))". However, the output shows that the minimum character count was applied to all counties. So, I searched on the web and found a way to use aggregate, which applies the function to the multiple columns.
<br>
Task 3 - It was challenging becuase I needed to put multiple functions and if statements all together in one function.
<br>
Task 4.1 - I was unable to use st_intersenct because of one geometry in the data. However, I found this code “sf::sf_use_s2(FALSE)” from the web, and the problem was solved. This sf_use_s2 was new to me, and I think I need more explanation about the code in addition to what the web explained: “sf_use_s2 returns the value of this variable before (re)setting it, invisibly if use_s2 is not missing.”
<br>
Additional information about sf_use_s2
<br>
Switching between S2 and GEOS - Edzer Pebesma and Dewey Dunnington
<br>
https://cran.r-project.org/web/packages/sf/vignettes/sf7.html
<br>
sf_use_s2(FALSE) can switch between s2Geometry and GEOS - Geometry Engine, Open Source which "The two-dimensional R2 library that was formerly used by sf." "GEOS treats projected coordinates as planar (i.e. two points lie on a line of infinite max lenght) while s2 is the two points lie on a great circle of circumference of 40 075 kilometers)."

### 3. What types of activities would you like to see in labs this semester?
<br>
I would like to do more labs creating animated plot maps with gganimate.