---
title: "Lab_5_Yoo"
author: "Kwang il Yoo"
date: "11/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Required Packages
library(sf)
library(tidyverse)
library(sp)
library(spdep)
library(raster)
library(leaflet)
library(RColorBrewer)
```

```{r}
#Prepare Data

#Lab 2
#County Shapefile
counties <- sf::read_sf("../data/CBW/County_Boundaries.shp") %>% sf::st_make_valid()
#BMP data
bmps <- read_csv("../data/CBW/BMPreport2016_landbmps.csv")

#Lab 3
#filtered States
states.filtered <- sf::read_sf("../data/states_filter.shp")

#Lab 4
#recent little wabash river shp file
lwr_new <- sf::read_sf("../data/LWR_Line.shp") %>% sf::st_make_valid()
#1940 little wabash river shp file
lwr_1940 <- sf::read_sf("../data/Left 1940.shp") %>% sf::st_make_valid()
##counties in IL shp file
IL_co <- sf::read_sf("../data/IL_BNDY_County_Py.shp") %>% sf::st_make_valid()
#DEM little wabash river
lwr_ras <- raster::raster("../data/lwr_ras.tif")
```

#Task 1
```{r}
#total cost of BMPs funded by county
bmps_trim <- bmps %>% mutate(., FIPS.trimmed = stringr::str_sub(GeographyName, 3, 5)) %>% group_by(FIPS.trimmed) %>% summarise(tcost = sum(Cost, na.rm = T))
#Join with county shp
counties_bmps <- left_join(counties, bmps_trim, by = c("COUNTYFP10" = "FIPS.trimmed"))
counties_bmps$tcost <- as.integer(counties_bmps$tcost)

#set basemap variable
Base_Map <- providers$CartoDB.Positron

#find equal interval breaks
c.max <- max(counties_bmps$tcost, na.rm = T)
c.min <- min(counties_bmps$tcost, na.rm = T)
br <- (c.max - c.min)/5
start <- 0
bins <- c()
for (i in 1:4) {
    start <- start+br
    bins <- append(bins, start)
  } 


#set bins and pal variables
pal <- colorBin(c("#F9E79F", "#7DCEA0", "#17A589", "#2E86C1", "#154360"), domain = counties_bmps$tcost, bins = c(0, bins, Inf), na.color = NA)

#map frame
base <- leaflet(counties_bmps) %>% 
      setView(lng = -76.55906218576634, lat = 40.507740442861356, zoom = 6) %>% addProviderTiles(Base_Map, group = "Base_Map")

#Hover Label
labels <- sprintf(
  "<strong>%s</strong><br/>$ %s total cost of BMPs",
  counties_bmps$NAME10, counties_bmps$tcost
) %>% lapply(htmltools::HTML)

#choropleth map
base %>%
      addPolygons(
        fillColor = ~pal(tcost),
        weight = 2,
        opacity = 1,
        color = "white",
        fillOpacity = 0.85,
        label = labels,
        labelOptions = labelOptions(
          direction = "bottom",
          style = list(
          "color" = "green",
          "font-family" = "serif",
          "font-style" = "italic",
          "box-shadow" = "3px 3px rgba(0,0,0,0.25)",
          "font-size" = "12px",
          "border-color" = "rgba(0,0,0,0.5)"
            )
          )
        ) %>%
      addLegend(
        pal = pal,
        values = ~tcost,
        title = "Total Cost of BMPs by County",
        position = "bottomright"
        )

```

#Task 2
```{r}
#Moran
states.filtered <- states.filtered %>% mutate(pct_b = (DP0080004/DP0080001)*100)
states.proj <- states.filtered %>% sf::st_transform(., "EPSG:4326")
states.q <- spdep::poly2nb(states.proj, queen = T)
states.w <- nb2listw(states.q, style="W")
black.pct.lag <- lag.listw(states.w, states.proj$pct_b)

#Local Moran
local <- localmoran(x = states.proj$pct_b, listw = states.w)
quadrant <- vector(mode="numeric",length=nrow(local))

# centers the variable of interest around its mean
m.qualification <- states.proj$pct_b - mean(states.proj$pct_b)     

# centers the local Moran's around the mean
m.local <- local[,] - mean(local[,1])    

# significance threshold
signif <- 0.1 

##mydata <- moran.plot() this will make dataframe for the moran including p values

# builds a data quadrant
quadrant[m.qualification >0 & m.local>0] <- 4  
quadrant[m.qualification <0 & m.local<0] <- 1      
quadrant[m.qualification <0 & m.local>0] <- 2
quadrant[m.qualification >0 & m.local<0] <- 3
quadrant[local[,5]>signif] <- 0   

# plot in r
lisa_col <- c("#F4F6F6", "blue", rgb(0,0,1,alpha=0.4), rgb(1,0,0,alpha=0.4), "red")
pal <- colorFactor(lisa_col, quadrant)


#set popup for p-value
p <- m.local[,5]
popup <- sprintf(
  "p-value: <strong>%g</strong>",
  p
) %>% lapply(htmltools::HTML)

#map it
base_moran <- leaflet(states.proj) %>% 
      setView(lng = -90.10238038443298, lat = 40.29523658379888, zoom = 6) %>%
      addProviderTiles(Base_Map, group = "Base_Map") %>%
      addProviderTiles(providers$Stamen.Toner, group = "Toner") %>%
      addProviderTiles(providers$Esri.NatGeoWorldMap, group = "ESRI")

base_moran %>%
      addPolygons(
        fillColor = ~pal(quadrant),
        weight = 0.5,
        opacity = 1,
        color = "gray",
        fillOpacity = 1,
        popup = popup
        ) %>%
      addLegend(
        colors = lisa_col,
        values = quadrant,
        title = "Black, African American Population Clustered",
        position = "bottomleft",
        labels = c("insignificant","low-low","low-high","high-low","high-high")
        ) %>%
      addLayersControl(
        baseGroups = c("Base_Map", "Toner", "ESRI"),
        options = layersControlOptions(collapsed = FALSE))

```

#Task 3
```{r}
#clip shapefile
clip_IL <- dplyr::filter(IL_co, COUNTY_NAM == "COLES" | COUNTY_NAM == "COLES" | COUNTY_NAM == "SHELBY" | COUNTY_NAM == "CUMBERLAND"| COUNTY_NAM == "EFFINGHAM" | COUNTY_NAM == "CLAY" | COUNTY_NAM == "RICHLAND" | COUNTY_NAM == "WAYNE" | COUNTY_NAM == "EDWARDS" | COUNTY_NAM == "WHITE" | COUNTY_NAM == "GALLATIN")

#check coordinate systems
crs(lwr_new)
crs(lwr_1940)
crs(lwr_ras)
crs(clip_IL)
#set coordinate system same as IL_co
st_crs(lwr_new) <- "EPSG:4326"
lwr_1940<- st_transform(lwr_1940, "EPSG:4326")
clip_IL <- st_transform(clip_IL, "EPSG:4326")

#Original Raster resolution 3x3 not work because of exceeding max memory use. (resolution set to 30x30 for faster loading time for the assignment)
#aggregate for less memory use (takes time...)
#lwr_ras <- raster::aggregate(lwr_ras, 10) <- save as new file
#writeRaster(lwr_ras, '../data/lwr_ras_agg.tif') <- saved
lwr_ras_agg <- raster::raster("../data/lwr_ras_agg.tif") #load agregated raster

#pal for ras
pal <- colorNumeric(c("#0C2C84", "#41B6C4", "#FFFFCC"), values(lwr_ras_agg),
  na.color = "transparent")

#map it!
base_lwr <- leaflet(clip_IL) %>% 
      setView(lng = -88.0551534406064, lat = 38.68125451362618, zoom = 8)%>%
      addProviderTiles(Base_Map, group = "Base_Map") %>%
      addProviderTiles(providers$Stamen.Toner, group = "Toner") %>%
      addProviderTiles(providers$Esri.NatGeoWorldMap, group = "ESRI")

base_lwr %>%
      addPolygons(
        data = clip_IL,
        color = "gray",
        weight = 0.5,
        opacity = 1,
        group = "counties",
        label = ~clip_IL$COUNTY_NAM
      ) %>%
      addPolylines(
        data = lwr_new,
        opacity = 0.5,
        weight = 1,
        color = "blue",
        group = "Stream Current"
        ) %>%
      addPolylines(
        data = lwr_1940,
        opacity = 0.5,
        weight = 1,
        color = "red",
        group = "Stream 1940"
        ) %>%
      addRasterImage(
        lwr_ras_agg,
        colors = pal,
        opacity = 0.8,
        ) %>%
      addLegend(
        values = 1,
        group = "Stream Current",
        position = "bottomleft",
        labels = "Stream Current",
        colors= "red"
        ) %>%
      addLegend(
        values = 2,
        group = "Stream 1940",
        position = "bottomleft",
        labels = "Stream 1940",
        colors= "blue"
        ) %>%
      addEasyButton(easyButton(
        icon="fa-globe", title="Zoom to Level 1",
        onClick=JS("function(btn, map){ map.setView([39.707181,-100.546875],4); }"))
        ) %>%
      addEasyButton(easyButton(
        icon="fa-home", title="Home",
        onClick=JS("function(btn, map){map.setView([38.68125451362618,-88.0551534406064],8); }"))
        ) %>%
      addEasyButton(easyButton(
        icon=htmltools::span(class = "1", htmltools::HTML("&#10112;")), title="Sector 1",
        onClick=JS("function(btn, map){map.setView([39.327924,-88.519805],10); }"))
        ) %>%
      addEasyButton(easyButton(
        icon=htmltools::span(class = "2", htmltools::HTML("&#10113;")), title="Sector 2",
        onClick=JS("function(btn, map){map.setView([38.610432,-88.324585],10); }"))
        ) %>%
      addEasyButton(easyButton(
        icon=htmltools::span(class = "3", htmltools::HTML("&#10114;")), title="Sector 3",
        onClick=JS("function(btn, map){map.setView([38.028622,-88.181763],10); }"))
        ) %>%
      addLayersControl(
        baseGroups = c("Base_Map", "Toner", "ESRI"),
        overlayGroups = c("counties", "Stream Current", "Stream 1940"),
        options = layersControlOptions(collapsed = FALSE)) %>%
      addMeasure()
```

#1. Reflect on the labs from this semester. What did you learn? What did you like? What did you not
like?

I like labs that have two parts. Such as lab 4, following the basic steps to make sure that I learned all the things that I need to have to create my own things. Those labs made me check whether I understood the contents correctly. Also, I was able to see whether I could apply those functions to my own project or custom creations.

Furthermore, I liked the bonus questions, not because of extra points but because there is a way to stretch my skills and knowledge without having pressure from getting full points.

I personally think that we should have more labs.

#2. Describe the “one thing” you chose to add to your map in Task 3 above. What did you do, and why
is it applicable to your map?
I added bookmark buttons to my map. Earth icon zoom out to US level so people can view the location of the study area. The home icon fits the extent of the study area. Icon 1, 2, 3 show the different sectors with a closer look. I think these bookmark buttons are very useful because my map needs to be zoomed in to view the detailed changes of the Little Wabash River.
